---
title: "covid 19"
author: "Elizabeth Rivera"
date: "12-October-2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Dying is not independent of sex.   
$H_A$: Men and women die at different rates, dying is dependent on sex.  


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
Among the females 25 were dead and 1831 were alive. Those among men, 44 were dead and 1411 were alive. According to the table above males have a higher death rate.  


## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this



  





  
## Numerical results





## Inferential results 




## Conclusions

