---
title: "covid 19"
author: "Elizabeth Rivera"
date: "12-October-2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Dying is independent of sex.   
$H_A$: Men and women die at different rates, dying is dependent on sex.  

The response variable is categorical and explanatory variable is categorial so we use the CAT-CAT helper file.  

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
Among the females 25 were dead and 1831 were alive. Those among men, 44 were dead and 1411 were alive. According to the table above males have a higher death rate.  


## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
Among those who died of COVID, the plot shows males are more likely die than females.   


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Based on the table above, it looks like twice the amount of males died compared to females.



## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the Chi squared the p-value is 0.0012 is less than 0.05 which means we reject the null hypothesis.  
In the Fisher Test the p-value is 0.0009 which means odds are males in Korea are 2.3 times more likely to die than females.   


## Conclusions

Among those in Korea males are most likely to die according to the Fisher test. The p-value is 0.0009. In the chi-squared test the p-value is 0.0012 which is lower than 0.05 so we reject the null hypothesis. Dying is not independent of sex. Males in Korea are more likely to die of COVID than females. 

